---
title: "Fuel_Consumption"
output: html_document
date: "2023-10-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(dplyr)
```
```{r}
data_1 <- read.csv('dataset.csv', na.strings = "-")
data_1
```
```{r}
missing_percentage <- colMeans(is.na(data_1))
missing_data <- data.frame(Column = names(data_1), MissingPercentage = missing_percentage)
print(missing_data)
```
```{r}
col_names_list <- list(colnames(data_1))
col_names_list[[1]]
#print(col_names_list)
```
```{r}
list_1 <- c("Torque.Nm", "CO..in.g.km..Average..g.km.", "Trip.Distance.km",
                 "Run.time.since.engine.start.s.", "Fuel.cost..trip..cost.", "Engine.Coolant.Temperature..C.",
                 "Litres.Per.100.Kilometer.Long.Term.Average..l.100km.", "Trip.Time.Since.journey.start..s.",
                 "Trip.time.whilst.moving..s.", "Trip.average.KPL.kpl.", "Fuel.Trim.Bank.1.Long.Term...",
                 "O2.Volts.Bank.1.sensor.1.V.", "O2.Volts.Bank.1.sensor.2.V.", "Fuel.flow.rate.hour.gal.hr.",
                 "Fuel.flow.rate.minute.gal.min.", "Fuel.Remaining..Calculated.from.vehicle.profile....",
                 "Horizontal.Dilution.of.Precision", "Acceleration.Sensor.Total..g.",
                 "Acceleration.Sensor.X.axis..g.", "Acceleration.Sensor.Y.axis..g.",
                 "Acceleration.Sensor.Z.axis..g.", "Speed..OBD..km.h.", "Trip.time.whilst.stationary..s.",
                 "Fuel_mass")
```
```{r}
list_2 <- c("Engine.kW..At.the.wheels..kW.", "Horsepower..At.the.wheels..hp.",
                 "Average.trip.speed.whilst.moving.only..km.h.", "Trip.average.Litres.100.KM.l.100km.",
                 "Average.trip.speed.whilst.stopped.or.moving..km.h.", "Absolute.Throttle.Position.B...","Relative.Throttle.Position...",
                 "Accelerator.PedalPosition.D...","Accelerator.PedalPosition.E...","Engine.Load.Absolute....",
                 "Fuel.trim.bank.1.sensor.1...","Fuel.Trim.Bank.1.Short.Term...",
                 "Throttle.Position.Manifold....","Timing.Advance...",
                 "Voltage..Control.Module..V.","Engine.Load...","Average.trip.speed.whilst.stopped.or.moving..km.h.","Kilometers.Per.Litre.Instant..kpl.",
                 "Miles.Per.Gallon.Instant..mpg.","Intake.Manifold.Pressure.psi.","Turbo.Boost...Vacuum.Gauge.psi.",
                 "Bearing","G.x.","G.y.","G.z.","G.calibrated.","Barometric.pressure..from.vehicle..psi.",
                 "Kilometers.Per.Litre.Long.Term.Average..kpl.","Miles.Per.Gallon.Long.Term.Average..mpg.",
                 "Engine.RPM.rpm.","Intake.Air.Temperature..C.")
```

###Parameters Discriptions
**Torque (Nm):** Torque is a measure of the engine's rotational force. It is an important factor in determining the engine's performance and efficiency.

**CO₂ in g/km (Average):** This column represents the average amount of carbon dioxide emissions produced per kilometer driven. It is a direct measure of the environmental impact of the engine's operation.

**Trip Distance (km):** This is the total distance covered during a specific trip.

**Run Time Since Engine Start (s):** It represents the duration for which the engine has been running since it was last started.

**Fuel Cost (Trip):** This column represents the cost of the fuel consumed during a specific trip.

**Engine Coolant Temperature (°C):** It indicates the temperature of the engine coolant, which is important for engine performance and efficiency.

**Litres Per 100 Kilometer (Long Term Average):** This is a measure of fuel efficiency, representing the amount of fuel (in liters) consumed per 100 kilometers traveled.

**Trip Time (Since Journey Start) (s):** It shows the total time elapsed since the start of the trip.

**Trip Time (Whilst Moving) (s):** This is the total time during the trip when the vehicle was in motion.

**Trip Average KPL (kpl):** This represents the average kilometers per liter of fuel consumed during the trip.

**Fuel Trim Bank 1 Long Term (%):** Fuel trim is an adjustment made to the air-fuel mixture. This parameter indicates a long-term adjustment percentage for Bank 1.

**O2 Volts Bank 1 Sensor 1 (V):** This is the voltage output of the first oxygen sensor in Bank 1.

**O2 Volts Bank 1 Sensor 2 (V):** Similar to the previous parameter, but for the second oxygen sensor in Bank 1.

**Fuel Flow Rate/Hour (gal/hr):** This represents the rate at which fuel is being consumed, measured in gallons per hour.

**Fuel Flow Rate/Minute (gal/min):** Similar to the previous parameter, but measured in gallons per minute.

**Fuel Remaining (Calculated from Vehicle Profile) (%):** This indicates the estimated remaining fuel percentage based on the vehicle's profile and usage.

**Horizontal Dilution of Precision:** This is a measure of the quality of the GPS signal, which can affect the accuracy of distance measurements.

**Acceleration Sensor (Total) (g):** It represents the total acceleration experienced by the vehicle, measured in 'g' units.

**Acceleration Sensor (X axis) (g):** This is the acceleration experienced along the X axis (usually lateral or side-to-side movement).

**Acceleration Sensor (Y axis) (g):** This is the acceleration experienced along the Y axis (usually longitudinal or front-to-back movement).

**Acceleration Sensor (Z axis) (g):** This is the acceleration experienced along the Z axis (usually vertical movement).

**Speed (OBD) (km/h):** This represents the vehicle's speed, as measured by the On-Board Diagnostics system.

**Trip Time (Whilst Stationary) (s):** This is the total time during the trip when the vehicle was stationary.

**Engine kW (At the wheels)(kW):** This represents the power output of the engine that actually reaches the wheels. It helps in understanding how efficiently the engine is converting fuel into usable power for the vehicle's movement.

**Horsepower (At the wheels)(hp):** Similar to engine kW, this measures the power at the wheels, but it's in horsepower units. It gives an alternative measurement of the engine's output.

**Average trip speed(whilst moving only)(km/h):** This indicates the average speed of the vehicle when it's in motion. It helps in understanding how speed affects fuel efficiency.

**Trip average Litres/100 KM(l/100km):** This shows how many liters of fuel are consumed to cover a distance of 100 kilometers. It's a direct measure of fuel efficiency.

**Average trip speed(whilst stopped or moving)(km/h):** This indicates the average speed of the vehicle including the time it's stopped. It helps in considering the effects of idling.

**Absolute Throttle Position B(%):** This is the percentage at which the throttle is open. It's important because it shows how much power the engine is producing relative to its maximum capacity.

**Relative Throttle Position(%):** This indicates the position of the throttle compared to its minimum and maximum limits. It helps in understanding how much power the engine is currently using.

**Accelerator Pedal Position D(%) and Accelerator Pedal Position E(%):** These show the percentage of depression of the accelerator pedal. It helps in understanding driver behavior and how aggressively the vehicle is being driven.

**Engine Load(Absolute)(%):** This represents the load on the engine, which is the amount of work it's doing. It's an important factor in fuel consumption because a heavily loaded engine may consume more fuel.

**Fuel trim bank 1 sensor 1(%) and Fuel Trim Bank 1 Short Term(%):** These values indicate adjustments made by the engine control unit to the air-fuel mixture. They are important for optimizing fuel consumption.

**Throttle Position(Manifold)(%):** This shows the position of the throttle in the manifold. It's important for understanding how the engine is controlling airflow, which affects fuel consumption.

**Timing Advance(°):** This represents the timing of the ignition events in the engine. Proper timing is crucial for efficient combustion and fuel consumption.

**Voltage (Control Module)(V):** This indicates the electrical voltage supplied to the control module. It's important for monitoring the electrical system, which indirectly affects fuel consumption.

**Kilometers Per Litre(Instant)(kpl) and Miles Per Gallon(Instant)(mpg):** These values directly indicate the vehicle's fuel efficiency in terms of distance covered per unit of fuel.

**Intake Manifold Pressure(psi) and Turbo Boost & Vacuum Gauge(psi):** These represent the pressure in the intake manifold and the boost pressure from a turbocharger. They are important for understanding how the engine is being fed air, which affects fuel consumption.

**Bearing, G(x), G(y), G(z), G(calibrated):** These likely refer to measurements related to acceleration and orientation of the vehicle. They can provide insight into driving behavior and conditions affecting fuel consumption.

**Barometric pressure (from vehicle)(psi):** This is the atmospheric pressure, which affects how much air the engine can take in. It's important for proper air-fuel mixture and combustion.

**Kilometers Per Litre(Long Term Average)(kpl) and Miles Per Gallon(Long Term Average)(mpg):** These values represent the long-term average fuel efficiency of the vehicle, which gives a more stable indication of its overall efficiency.

**Engine RPM(rpm):** This indicates how fast the engine is spinning. It's crucial for understanding engine performance and fuel consumption, as higher RPMs often mean higher fuel consumption.

**Intake Air Temperature(°C):** This measures the temperature of the air entering the engine. Cooler air is denser, which can affect combustion efficiency and fuel consumption.


```{r}
library(ggplot2)
# Define the label
label <- 'Fuel_mass'

# Create a histogram using ggplot
ggplot(data_1, aes(x = as.numeric(ifelse(is.na(data_1[[label]]), 0, as.character(data_1[[label]]))))) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(x = label, y = "Frequency") +
  theme_minimal()

# Print the minimum and maximum values
cat("Minimum Value:", min(as.numeric(ifelse(is.na(data_1[[label]]), 0, as.character(data_1[[label]])))), "\n")
cat("Maximum Value:", max(as.numeric(ifelse(is.na(data_1[[label]]), 0, as.character(data_1[[label]])))), "\n")
```
**From the above plot we can see that the distribution has a long right tail, which means that there are relatively few extreme values or outliers on the high end of fuel consumption. These outliers represent instances where fuel consumption is significantly higher than the majority of observations.**

```{r}
# library
library(ggplot2)
library(dplyr)
library(hrbrthemes)

# Start with your dataset 'datset'
datset <- data_1
varaiable <- "Torque.Nm."
p <- datset %>%
  mutate(!!sym(varaiable) := ifelse(is.na(!!sym(varaiable)), 0, !!sym(varaiable))) %>%
  mutate(bin = cut(!!sym(varaiable), breaks = 4)) %>%
  ggplot(aes(x = bin, y = Fuel_mass)) +
  geom_boxplot(fill = "blue") +
  theme_minimal() +
  xlab("Torque.Nm. Bins")

# Show the plot
print(p)
```

```{r}
# Create a vector containing the column names want to process
trimmable_parameters <- c("Torque.Nm.", "CO..in.g.km..Average..g.km.", "Trip.Distance.km.", "Run.time.since.engine.start.s.", "Fuel.cost..trip..cost.", "Engine.Coolant.Temperature..C.", "Litres.Per.100.Kilometer.Long.Term.Average..l.100km.", "Trip.Time.Since.journey.start..s.", "Trip.time.whilst.moving..s.", "Trip.average.KPL.kpl.", "Fuel.Trim.Bank.1.Long.Term...", "O2.Volts.Bank.1.sensor.1.V.", "O2.Volts.Bank.1.sensor.2.V.", "Fuel.flow.rate.hour.gal.hr.", "Fuel.flow.rate.minute.gal.min.", "Fuel.Remaining..Calculated.from.vehicle.profile....", "Horizontal.Dilution.of.Precision", "Acceleration.Sensor.Total..g.", "Acceleration.Sensor.X.axis..g.", "Acceleration.Sensor.Y.axis..g.", "Acceleration.Sensor.Z.axis..g.", "Speed..OBD..km.h.", "Trip.time.whilst.stationary..s.")


# Create an empty list to store the plots
plot_list <- list()
dataset <- data_1
# Loop through the vector of column names
for (variable in trimmable_parameters) {
  
  # Replace NA with zero in the dataset for the current column
  dataset <- dataset %>%
    mutate(!!sym(variable) := ifelse(is.na(!!sym(variable)), 0, !!sym(variable))) 
  
  # Create bins for the current column
  dataset <- dataset %>%
    mutate(bin = cut(!!sym(variable), breaks = 4))
  
  # Create a boxplot for the current column
  p <- ggplot(data = dataset, aes(x = bin, y = Fuel_mass)) +
    geom_boxplot(fill = "#69b3a2") +
    theme_minimal() +
    xlab(variable)
  
  # Store the plot in the list
  plot_list[[variable]] <- p
}

# Print or display the plots from the list
for (i in 1:length(plot_list)) {
  print(plot_list[[i]])
}
```

For the below columns, convert time values in seconds to minutes for the following columns:
1) Run.time.since.engine.start.s.
2) Trip.Time.Since.journey.start..s.
3) Trip.time.whilst.moving..s.
4) Trip.time.whilst.stationary..s.

```{r}
data_1 <- data_1 %>%
  mutate(Run.time.since.engine.start.s.mins = Run.time.since.engine.start.s. / 60)

data_1 <- data_1 %>%
  mutate(Trip.Time.Since.journey.start..s.mins = Trip.Time.Since.journey.start..s. / 60)

data_1 <- data_1 %>%
  mutate(Trip.time.whilst.moving..s.mins = Trip.time.whilst.moving..s. / 60)

data_1 <- data_1 %>%
  mutate(Trip.time.whilst.stationary..s.mins = Trip.time.whilst.stationary..s. / 60)

```

```{r}
# Create a vector containing the column names you want to process
Trip_sec_to_minutes<- c("Run.time.since.engine.start.s.mins","Trip.Time.Since.journey.start..s.mins","Trip.time.whilst.moving..s.mins","Trip.time.whilst.stationary..s.mins")

# Create an empty list to store the plots
plot_list <- list()
dataset <- data_1
# Loop through the vector of column names
for (variable in Trip_sec_to_minutes) {
  
  # Replace NA with zero in the dataset for the current column
  dataset <- dataset %>%
    mutate(!!sym(variable) := ifelse(is.na(!!sym(variable)), 0, !!sym(variable))) 
  
  # Create bins for the current column
  dataset <- dataset %>%
    mutate(bin = cut(!!sym(variable), breaks = 4))
  
  # Create a boxplot for the current column
  p <- ggplot(data = dataset, aes(x = bin, y = Fuel_mass)) +
    geom_boxplot(fill = "#69b3a2") +
    theme_minimal() +
    xlab(variable)
  
  # Store the plot in the list
  plot_list[[variable]] <- p
}

# Print or display the plots from the list
for (i in 1:length(plot_list)) {
  print(plot_list[[i]])
}
```
