---
title: "Voltage Related Columns"
output: html_document
date: "2023-10-24"
---
```{r}
data1=read.csv("C:/Users/KDIVA/OneDrive/Desktop/Intro_to_data_science_project/dataset.csv")
colnames(data1)
```
```{r}
data1$Cluster=as.factor(data1$Cluster)
library(ggplot2)
ggplot(data1, aes(x = Fuel_mass, y = Cluster)) +
  geom_boxplot(fill = "#66c2a5", color = "#7570b3") +
  labs(title = "Fuel Mass Distribution by Cluster", x = "Fuel Mass", y = "Cluster")
```



Clusters- 1,2 and 3 are similar. The analysis would be done such that the driver should avoid cluster 5 as it has the highest fuel mass consumption. 



O2 Volts Bank 1 Sensor 1 (V): 

This is the voltage output of the first oxygen sensor in Bank 1.


W.R.T Fuel Consumption: 

This sensor provides feedback on the air-fuel mixture. Proper mixture leads to better fuel economy. The mixture should not be too lean or too rich.



```{r}
ggplot(data1, aes(x = O2.Volts.Bank.1.sensor.1.V., y = Cluster)) +
  geom_boxplot(fill = "#66c2a5", color = "#7570b3") +
  labs(title = "O2 Volts Bank 1 Sensor 1 (V) Distribution by Cluster", x = "O2 Volts Bank 1 Sensor 1 (V)", y = "Cluster")
```
```{r}


ci_list <- list()

for (cluster_id in unique(data1$Cluster)) {
  subset_data <- data1[data1$Cluster == cluster_id, ]
  ci <- t.test(subset_data$O2.Volts.Bank.1.sensor.1.V.)
  ci_list[[as.character(cluster_id)]] <- ci$conf.int
}

print(ci_list)

```
```{r}
data_o2_sen1 <- data1[data1$Cluster == "4" | data1$Cluster == "5",]
ggplot(data_o2_sen1, aes(x = O2.Volts.Bank.1.sensor.1.V., fill = Cluster)) +
  geom_histogram(position = "identity", alpha = 0.4)

```



There is no clear right or left shift. So, we bifercated the column into greater than 0.45 and less than 0.45 to do further analysis.



```{r}

less_than_045 <- subset(data1, O2.Volts.Bank.1.sensor.1.V. < 0.45)



ci_list <- list()

for (cluster_id in unique(less_than_045$Cluster)) {
  subset_data <- less_than_045[less_than_045$Cluster == cluster_id, ]
  ci <- t.test(subset_data$O2.Volts.Bank.1.sensor.1.V.)
  ci_list[[as.character(cluster_id)]] <- ci$conf.int
}

print(ci_list)
```
```{r}
data_o2_lower <- less_than_045[less_than_045$Cluster == "1" | less_than_045$Cluster == "5",]
ggplot(data_o2_lower, aes(x = O2.Volts.Bank.1.sensor.1.V., fill = Cluster)) +
  geom_histogram(position = "identity", alpha = 0.4)


```



Observation:

There is a right shift from cluster 5[C.I -0.10 to 0.13] to cluster 1[C.I -0.11 to 0.15]




```{r}
greater_than_045 <- subset(data1, O2.Volts.Bank.1.sensor.1.V. > 0.45)



ci_list <- list()

for (cluster_id in unique(greater_than_045$Cluster)) {
  subset_data <- greater_than_045[greater_than_045$Cluster == cluster_id, ]
  ci <- t.test(subset_data$O2.Volts.Bank.1.sensor.1.V.)
  ci_list[[as.character(cluster_id)]] <- ci$conf.int
}

print(ci_list)
```
```{r}
data_o2_upper <- greater_than_045[greater_than_045$Cluster == "4" | greater_than_045$Cluster == "5",]
ggplot(data_o2_upper, aes(x = O2.Volts.Bank.1.sensor.1.V., fill = Cluster)) +
  geom_histogram(position = "identity", alpha = 0.4)

```




Observation:

There is a left shift from cluster 5[C.I -0.80 to 0.83] to cluster 4[C.I -0.79 to 0.81]. We can say that the column O2 Volts Bank 1 Sensor 1 is not a signifiant factor which affects Fuel consumption.